---
output:
  pdf_document: default
  html_document: default
---
# Set up
```{r setup, include = FALSE}

# Load required packages
library(dplyr)
library(broom)
library(modelr)

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(infer))
```

# 1. Intro & DataSet
```{r}
# Read three csv files
emissions <- read.csv("emissions.csv")
save_the_bees <- read.csv("save_the_bees.csv")
us_tornado_1950_2021 <-
  read.csv("us_tornado_dataset_1950_2021.csv")
```

```{r}
# Filter year 2015-2021 to use same period
emission_21 <-filter(emissions, year >= 2015 & year <= 2021)
bee_21 <- filter(save_the_bees, year >= 2015 & year <= 2021)
tornado_21 <- filter(us_tornado_1950_2021, yr >= 2015 & yr <= 2021)

```

### Honey Bee data
```{r}
# Select useful data to calculate average max_colonies with quarter
bee_avg_max <- bee_21 %>%
  select( state, state_code, quarter, year, max_colonies)
```

```{r}
# Pivot wider into q1~q4 
bee_avg_max <- bee_avg_max %>%
  pivot_wider(
    names_from = quarter,
    values_from = max_colonies
  )

bee_avg_max <- bee_avg_max %>%
  rename(q1 = '1', q2 = '2', q3 = '3', q4 = '4')
```

```{r}
# Mutate average max colonies
bee_avg_max <- bee_avg_max %>%
  mutate(
    avg_max_col = if_else(
      year != "2019",
      (q1 + q2 + q3 + q4) / 4, (q1 + q3 + q4) / 3) )
```

### CO2 emission
```{r}
# Sum of each fuel.name values
emission_fuel <- emission_21 %>%
  group_by(year,state.name, fuel.name) %>%
  summarise(sum_value = sum(value, na.rm = TRUE),.groups = 'drop')

```

```{r}
# Pivot wide each fuel name
emission_fuel <- emission_fuel %>%
  pivot_wider(
    names_from = fuel.name,
    values_from = sum_value
  )%>%
  rename(state = state.name, all_fuel = 'All Fuels')
```

### Tornado
```{r}
# Count tornado frequency by year and state
tornado_freq <- tornado_21 %>%
  rename(year = yr, state = st) %>%
  group_by(year, state) %>%
  summarise(count = n(), .groups = 'drop') 
```

```{r}
# Fill 0 at no tornado state&year 
# Make same column number with emission_21
tornado_freq <- tornado_freq %>%
  complete(year = unique(tornado_freq$year),
           state = unique(tornado_freq$state),
           fill = list(count = 0)) ### fill 0
```

### Merge Data for Modeling

## 4. Modeling 1: Make first dataframe and model

i. Step 1: bee vs emission
```{r}
# Data cleaning
bee_state <- bee_avg_max %>%
  filter(!(state %in% c("Other", "United States")))
emission_state <- emission_fuel %>%
  filter(!(state %in% c("Alaska","Delaware","District of Columbia",
                        "Nevada","New Hampshire","Rhode Island",
                        "United States")))
```

```{r}
# Data merge
bee_emission_df <- merge(bee_state, emission_state, by = c("state", "year")) %>%
  mutate(
    avg_max_col = round(avg_max_col, 0),
    all_fuel = round(all_fuel, 0)
    )
```


ii. Step 2: bee vs tornado
```{r}
# Data cleaning
tornado_state <- tornado_freq %>%
  filter(!(state %in% c("DC", "DE","NH","NV","PR","RI","VI"))
         ) %>%
  rename(state_code = state)
```

```{r}
# Data merge
bee_tornado_df <- merge(bee_state, tornado_state, by = c("state_code", "year"))
```


iii. Step 3: bee vs emission + tornado
```{r}
# Data cleaning
bee_tornado_df <- bee_tornado_df %>%
  select(year, state, count)
bee_emission_df <- bee_emission_df %>%
  select(year, state, avg_max_col, all_fuel)
```

```{r}
# Data merge
bee_df <- merge(bee_emission_df, bee_tornado_df, by = c("state", "year"))
```

------------------------------------------------------------------------

## 2. EDA / Summary Statistics

i. Data frame
```{r}
bee_df %>%
  ggplot() +
  geom_point(mapping = aes(x = year, y = avg_max_col)) +
  facet_wrap(~state, scales = "free_y")+
  geom_smooth(aes(year, avg_max_col),method = "lm")
```

```{r}
new_bee_df <- bee_df %>%
  filter(state %in% c("Arizona", "Arkansas", "Colorado", "Louisiana", 
                      "Maine", "Maryland", "Minnesota", "Mississippi",
                      "Missouri", "Montana", "Nebraska", "New Mexico",
                      "Ohio", "South Dakota", "Washington", "Wisconsin"))
```

ii. Summary statistics
```{r}
#Dividing avg_max_col by 1000
new_bee_df <- new_bee_df %>%
  mutate(avg_max_col_1000 = avg_max_col/1000)
```

```{r}
# Detailed summary statistics for the entire dataset
avg_max_col_summary <- new_bee_df %>%
  summarise(
    avg_max_col_mean = mean(avg_max_col_1000, na.rm = TRUE),
    avg_max_col_sd = sd(avg_max_col_1000, na.rm = TRUE),
    avg_max_col_min = min(avg_max_col_1000, na.rm = TRUE),
    avg_max_col_max = max(avg_max_col_1000, na.rm = TRUE))

all_fuel_summary <- new_bee_df %>%
  summarise(
    all_fuels_mean = mean(all_fuel, na.rm = TRUE),
    all_fuels_sd = sd(all_fuel, na.rm = TRUE),
    all_fuels_min = min(all_fuel, na.rm = TRUE),
    all_fuels_max = max(all_fuel, na.rm = TRUE))

tornado_summary <- new_bee_df %>%
  summarise(
    tornado_count_mean = mean(count, na.rm = TRUE),
    tornado_count_sd = sd(count, na.rm = TRUE),
    tornado_count_min = min(count, na.rm = TRUE),
    tornado_count_max = max(count, na.rm = TRUE))
```

```{r}
print(avg_max_col_summary)
```

```{r}
print(all_fuel_summary)
```

```{r}
print(tornado_summary)
```

----------------------------------------------------------------------------

iii. Visualization
```{r}
# Visualization of mean average maximum colonies by state and year
new_bee_df %>%
  group_by(year, state) %>%
  summarise(
    avg_max_col_mean = mean(avg_max_col_1000, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  ggplot() +
  geom_bar(mapping = aes(x = state, y = avg_max_col_mean, fill = year), 
           stat = "identity", position = "dodge") +
  labs(
    title = "Average Maximum Colonies by State and Year",
    x = "State",
    y = "Average Maximum Colonies (1 = 1000)"
  ) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
#Histogram of average bees
new_bee_df %>%
  ggplot() +
  geom_histogram(mapping = aes(x = avg_max_col_1000,fill = state), bins = 20) +
  facet_wrap(~year) +
  labs(title = "Distribution of Average Maximum Colonies",
       x = "Average Maximum Colonies (1 = 1000)",
       y = "Frequency")
```
```{r}
#Histogram of Carbon Emission
new_bee_df %>%
  ggplot() +
  geom_histogram(mapping = aes(x = all_fuel,fill = state), bins = 20) +
  facet_wrap(~year) +
  labs(title = "Distribution of Carbon Emissions",
       x = "Carbon Emission",
       y = "Frequency")
```
```{r}
#Histogram of Tornado
new_bee_df %>%
  ggplot() +
  geom_histogram(mapping = aes(x = count, fill = state), bins = 20) +
  facet_wrap(~year) +
  labs(title = "Distribution of Tornado",
       x = "Carbon Emission",
       y = "Frequency")
```
```{r,}
#Boxplot of average colonies
new_bee_df %>%
  ggplot() +
  geom_boxplot(mapping = aes(x = avg_max_col_1000), bins = 20) +
  facet_wrap(~year) +
  labs(title = "Distribution of Average Maximum Colonies",
       x = "Average Maximum Colonies (1 = 1000)",
       y = "Frequency")
```
```{r,}
#Boxplot of Emission
new_bee_df %>%
  ggplot() +
  geom_boxplot(mapping = aes(x = all_fuel), bins = 20) +
  facet_wrap(~year) +
  labs(title = "Distribution of Carbon Emission",
       x = "Carbon Emission",
       y = "Frequency")
```
```{r,}
#Boxplot of Tornado
new_bee_df %>%
  ggplot() +
  geom_boxplot(mapping = aes(x = count), bins = 20) +
  facet_wrap(~year) +
  labs(title = "Distribution of Tornado",
       x = "Count of tornado",
       y = "Frequency")
```
```{r}
#Scatter plot for Colonies and Emission
new_bee_df %>%
  ggplot() +
  geom_point(mapping = aes(x = all_fuel,y = avg_max_col_1000)) +
  geom_smooth(aes(all_fuel,avg_max_col_1000),method = "lm")+
  facet_wrap(~year, scales = "free")+
  labs(
    title = "Distribution of Average Maximum Colonies and Emission by Year",
    x = "All fuels",
    y = "Maximum Colonies (1 = 1000)"
    )
```
```{r}
#Violin plot for Emission by year
new_bee_df %>%
  ggplot() +
  geom_violin(mapping = aes(x = factor(year), y = all_fuel)) +
  labs(
    title = "Distribution of Carbon Emissions by Year",
    x = "Year",
    y = "Amount of Emissions"
    )
```
```{r}
#Violin plot for Tornado count
new_bee_df %>%
  ggplot() +
  geom_violin(mapping = aes(x = factor(year),y = count)) +
  labs(
    title = "Distribution of Tornado Count by Year",
    x = "Year",
    y = "Tornado count"
    )
```


-----------------------------------------------------------------------


## 3. Hypothesis testing Idea

-----------------------------------------------------------------------


iv. Modeling: Multiple linear regression

## 4. Modeling 1: Make first dataframe(1.Data) and model
```{r}
# Modeling
bee_model_0 <- lm(avg_max_col ~ all_fuel + count + all_fuel*count, data = bee_df)
```

* Step 1: Examining the model coefficients & model fit
```{r}
# tidy() -> y-intercept , lm-slope
# glance() -> R2 value
bee_model_0 %>%
  tidy()

bee_model_0 %>%
  glance() %>%
  select(1:3)
```
* Step 2: Residuals
```{r}
# add predictions and residuals
bee_df <- bee_df %>%
add_predictions(bee_model_0) %>%
add_residuals(bee_model_0)
```

* Step 3: 3 assumptions
```{r}
# Linearity: Observed vs Predicted Plot
bee_df %>%
ggplot()+
geom_point(mapping = aes(x = pred, y = avg_max_col))+
geom_abline(slope = 1, intercept = 0, color = "blue")+
labs(title = "Observed Predictions of factors",
x= "Predictions", y = "Observed Price")
```

```{r}
# Nearly normal residuals: QQ plot
bee_df %>%
ggplot()+
geom_qq(aes(sample = resid))+
geom_qq_line(aes(sample = resid),color = "red")+
labs(title = "Q-Q Plot showing Residuals of Each Data",
x = "Theoretical Quantiles", y = "Sample Quantiles")
```

```{r}
# Constant variability: a residuals vs predicted plot
bee_df %>%
ggplot()+
geom_point(mapping = aes(x = pred, y = resid))+
geom_hline(yintercept = 0)+
labs(title = "Residual Versus Predicted of factors",
x = "Predicted", y = "Residual")
```

## Modeling 2: Making New Dataframe using decreasing lm state

```{r}
# Examine data
bee_df %>%
  ggplot() +
  geom_point(mapping = aes(x = year, y = avg_max_col)) +
  facet_wrap(~state, scales = "free_y")+
  geom_smooth(aes(year, avg_max_col),method = "lm")
```

```{r}
# Filter states showing negative linear relationship
new_bee_df <- bee_df %>%
  filter(state %in% c("Arizona", "Arkansas", "Colorado", "Louisiana", 
                      "Maine", "Maryland", "Minnesota", "Mississippi",
                      "Missouri", "Montana", "Nebraska", "New Mexico",
                      "Ohio", "South Dakota", "Washington", "Wisconsin"))
```

```{r}
# Make new dataframe
new_bee_df <- new_bee_df %>%
add_predictions(bee_model_0) %>%
add_residuals(bee_model_0)
```

```{r}
# Linearity: with new df
new_bee_df %>%
ggplot()+
geom_point(mapping = aes(x = pred, y = avg_max_col))+
geom_abline(slope = 1, intercept = 0, color = "blue")+
labs(title = "Observed Predictions of factors",
x= "Predictions", y = "Observed Price")
```

```{r}
# Nearly normal residuals: with new df
new_bee_df %>%
ggplot()+
geom_qq(aes(sample = resid))+
geom_qq_line(aes(sample = resid),color = "red")+
labs(title = "Q-Q Plot showing Residuals of Each Data",
x = "Theoretical Quantiles", y = "Sample Quantiles")
```

```{r}
# Constant variability: with new df
new_bee_df %>%
ggplot()+
geom_point(mapping = aes(x = pred, y = resid))+
geom_hline(yintercept = 0)+
labs(title = "Residual Versus Predicted of factors",
x = "Predicted", y = "Residual")
```



## Modeling 3: new_avg_max_col = avg_max_col / 1000

```{r}
new_bee_df_02<- new_bee_df %>%
  mutate(new_avg_max_col = avg_max_col / 1000)
```

...


## Modeling 4: Make stronger model by adding Internal Factors 'varroa_mites'

i. Make new dataset
```{r}
bee_inf <- filter(save_the_bees, year >= 2015 & year <= 2021)

bee_inf <- bee_inf %>%
  select( state, state_code, quarter, year, varroa_mites)
```

```{r}
bee_inf <- bee_inf %>%
  pivot_wider(
    names_from = quarter,
    values_from = varroa_mites
  )
```

```{r}
bee_inf <- bee_inf %>%
  rename(q1 = '1', q2 = '2', q3 = '3', q4 = '4')
```


```{r}
# Mtate average varroa_mites
bee_inf <- bee_inf %>%
  mutate(
    avg_var_col = if_else(
      year != "2019",
      (q1 + q2 + q3 + q4) / 4, (q1 + q3 + q4) / 3) )
```

```{r}
bee_inf <- bee_inf %>%
  filter(!(state %in% c("Other", "United States")))
```

ii. Make new dataframe
```{r}
bee_df_2 <- merge(bee_df, bee_inf, by = c("state", "year"))
```

iii. Make new model (ver.2)
```{r}
new_bee_model <- lm(avg_max_col ~ all_fuel + count + avg_var_col + all_fuel*count,  data = bee_df_2)
```

iv. Examining new model -1 
```{r}
# tidy(), glance()
new_bee_model %>%
  tidy()

new_bee_model %>%
  glance() %>%
  select(1:3)
```

v. Examining new model -2
```{r}
new2_bee_df <- bee_df_2 %>%
add_predictions(new_bee_model) %>%
add_residuals(new_bee_model)
```

Vi. Filter states again
```{r}
new2_bee_df <- new2_bee_df %>%
  filter(state %in% c("Arizona", "Arkansas", "Colorado", "Louisiana", 
                      "Maine", "Maryland", "Minnesota", "Mississippi",
                      "Missouri", "Montana", "Nebraska", "New Mexico",
                      "Ohio", "South Dakota", "Washington", "Wisconsin"))
```


```{r}
# Linearity: with new df and new model
new2_bee_df %>%
ggplot()+
geom_point(mapping = aes(x = pred, y = avg_max_col))+
geom_abline(slope = 1, intercept = 0, color = "blue")+
labs(title = "Observed Predictions of factors",
x= "Predictions", y = "Observed Price")
```

```{r}
# Nearly normal residuals: with new df and new model
new2_bee_df %>%
ggplot()+
geom_qq(aes(sample = resid))+
geom_qq_line(aes(sample = resid),color = "red")+
labs(title = "Q-Q Plot showing Residuals of Each Data",
x = "Theoretical Quantiles", y = "Sample Quantiles")
```

```{r}
# Constant variability: with new df and new model
new2_bee_df %>%
ggplot()+
geom_point(mapping = aes(x = pred, y = resid))+
geom_hline(yintercept = 0)+
labs(title = "Residual Versus Predicted of factors",
x = "Predicted", y = "Residual")
```


