---
output:
  pdf_document: default
  html_document: default
---
#Set up
```{r setup, include = FALSE}

# Load required packages
library(dplyr)
library(broom)
library(modelr)

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(infer))
```

#[Intro & DataSet]

```{r}
emissions <- read.csv("emissions.csv")
save_the_bees <- read.csv("save_the_bees.csv")
us_tornado_1950_2021 <-
  read.csv("us_tornado_dataset_1950_2021.csv")
```

### filter 2015-2021 
```{r}
emission_21 <-filter(emissions, year >= 2015 & year <= 2021)
bee_21 <- filter(save_the_bees, year >= 2015 & year <= 2021)
tornado_21 <- filter(us_tornado_1950_2021, yr >= 2015 & yr <= 2021)

```

## [Honey Bee]

### select useful data 
```{r}
bee_21 <- bee_21 %>%
  select( state, state_code, quarter, year, max_colonies)
```

### pivot wider into q1~q4 
```{r}
bee_21 <- bee_21 %>%
  pivot_wider(
    names_from = quarter,
    values_from = max_colonies
  )
```


```{r}
bee_21 <- bee_21 %>%
  rename(q1 = '1', q2 = '2', q3 = '3', q4 = '4')
```

### mutate average max colonies
```{r}
bee_21 <- bee_21 %>%
  mutate(
    avg_max_col = if_else(
      year != "2019",
      (q1 + q2 + q3 + q4) / 4, (q1 + q3 + q4) / 3) )
```




## [CO2 emission]

### sum of each fuel.name value
```{r}
emission_21 <- emission_21 %>%
  group_by(year,state.name, fuel.name) %>%
  summarise(sum_value = sum(value, na.rm = TRUE),.groups = 'drop')

```
### pivot wide each fuel name
```{r}
emission_21 <- emission_21 %>%
  pivot_wider(
    names_from = fuel.name,
    values_from = sum_value
  )%>%
  rename(state = state.name, all_fuel = 'All Fuels')
```





## [Tornado]

### Frequency

```{r}
tornado_21 <- tornado_21 %>%
  group_by(yr, st) %>%
  summarise(
    count= n(), .groups = 'drop'
  ) %>%
  rename(year = yr, state = st)
```

```{r}
tornado_10 <- tornado_21 %>%
  filter(state %in% c("ND", "CA", "TX", "MT", "FL", "SD", "MN", "OR", "MI", "GA")) %>%
  mutate(state = recode(state, 
                     "ND" = "North Dakota", "CA" = "California",
                     "TX" = "Texas", "MT" = "Montana",
                     "FL" = "Florida","SD" = "South Dakota",
                     "MN" = "Minnesota","OR" = "Oregon",
                     "MI" = "Michigan","GA" = "Georgia"))
```


## EDA / Summary Statistics

```{r}
#Dividing avg_max_col by 1000
bee_df <- bee_df %>%
  mutate(avg_max_col_1000 = avg_max_col/1000)
```

```{r}
# Detailed summary statistics for the entire dataset
avg_max_col_summary <- bee_df %>%
  summarise(
    avg_max_col_mean = mean(avg_max_col_1000, na.rm = TRUE),
    avg_max_col_sd = sd(avg_max_col_1000, na.rm = TRUE),
    avg_max_col_min = min(avg_max_col_1000, na.rm = TRUE),
    avg_max_col_max = max(avg_max_col_1000, na.rm = TRUE))

all_fules_summary <- bee_df %>%
  summarise(
    all_fuels_mean = mean(all_fules, na.rm = TRUE),
    all_fuels_sd = sd(all_fules, na.rm = TRUE),
    all_fuels_min = min(all_fules, na.rm = TRUE),
    all_fuels_max = max(all_fules, na.rm = TRUE))

tornado_summary <- bee_df %>%
  summarise(
    tornado_count_mean = mean(count, na.rm = TRUE),
    tornado_count_sd = sd(count, na.rm = TRUE),
    tornado_count_min = min(count, na.rm = TRUE),
    tornado_count_max = max(count, na.rm = TRUE))
```

```{r}
print(avg_max_col_summary)
```

```{r}
print(all_fules_summary)
```

```{r}
print(tornado_summary)
```

## [EDA / Visualization]





## [Hypothesis testing Idea]


## [Modeling ] (lm)

1. bee vs emission

### data cleaning
```{r}
bee_state <- bee_21 %>%
  filter(!(state %in% c("Other", "United States")))
emission_state <- emission_21 %>%
  filter(!(state %in% c("Alaska","Delaware","District of Columbia",
                        "Nevada","New Hampshire","Rhode Island",
                        "United States")))
```

### data merge
```{r}
bee_emission_df <- merge(bee_state, emission_state, by = c("state", "year")) %>%
  mutate(
    avg_max_col = round(avg_max_col, 0),
    all_fuel = round(all_fuel, 0)
    )
```


2. bee vs tornado

### data cleaning
```{r}
tornado_state <- tornado_21 %>%
  filter(!(state %in% c("DC", "DE","NH","NV","PR","RI","VI"))
         ) %>%
  rename(state_code = state)
```

### data merge
```{r}
bee_tornado_df <- merge(bee_state, tornado_state, by = c("state_code", "year"))
```


3. bee vs emission + tornado

### data cleaning
```{r}
bee_tornado_df <- bee_tornado_df %>%
  select(year, state, count)
bee_emission_df <- bee_emission_df %>%
  select(year, state, avg_max_col, all_fuel)
```

### data merge
```{r}
bee_df <- merge(bee_emission_df, bee_tornado_df, by = c("state", "year"))
```


4. modeling _multiple linear regression

### modeling
```{r}
bee_model_0 <- lm(avg_max_col ~ all_fuel + count + all_fuel*count, data = bee_df)
```

### tidy(), glance()
```{r}
bee_model_0 %>%
  tidy()

bee_model_0 %>%
  glance() %>%
  select(1:3)
```

### add predictions and residuals
```{r}
bee_df <- bee_df %>%
add_predictions(bee_model_0) %>%
add_residuals(bee_model_0)
```

5. modeling _ graphing

### Linearity: Observed vs Predicted Plot
```{r}
bee_df %>%
ggplot()+
geom_point(mapping = aes(x = pred, y = avg_max_col))+
geom_abline(slope = 1, intercept = 0, color = "blue")+
labs(title = "Observed Predictions of factors",
x= "Predictions", y = "Observed Price")
```

### Nearly normal residuals: QQ plot
```{r}
bee_df %>%
ggplot()+
geom_qq(aes(sample = resid))+
geom_qq_line(aes(sample = resid),color = "red")+
labs(title = "Q-Q Plot showing Residuals of Each Data",
x = "Theoretical Quantiles", y = "Sample Quantiles")
```

### Constant variability: a residuals vs predicted plot
```{r}
bee_df %>%
ggplot()+
geom_point(mapping = aes(x = pred, y = resid))+
geom_hline(yintercept = 0)+
labs(title = "Residual Versus Predicted of factors",
x = "Predicted", y = "Residual")
```






